---
title: "Clustering Twitter Users"
output:
  html_document:
    df_print: paged
---

A quick and dirty look at trying to find inauthentic (bot) behavior in Twitter using graph clustering.

```{r, message=FALSE}
# load libraries to make this easier
library(stringdist)
library(visNetwork)
library(dplyr)
library(igraph)
library(data.table)
library(rtweet)
library(lubridate)
```

Look at tweets around "dictatorBiden" - Note: assumes you've already set up rtweet according to the instructions [here](https://cran.r-project.org/web/packages/rtweet/vignettes/auth.html)

```{r, eval=FALSE}
q<-'dictatorbiden'
rt<-search_tweets(q,n=10000)
setDT(rt) # make this into a data.table for easy handling (could do all of this in tidyverse, as well)
```

```{r, echo=FALSE}
# load the dataset collected previously on 9/11/21
rt<-fread('dictatorbiden_tweets.gz')
```

Now with a data table of tweets, subset it to get a unique frame of IDs spreading these tweets. Keep the created date so we can see another element of potential inauthentic behavior - coincident acount creations beyond what would be expected in the "birthday paradox"

```{r}
# get just the date
rt[,acct_date:=date(account_created_at)]
# create a subset a where we count the unique number
a<-unique(rt[,.(screen_name,acct_date)])
a[,N:=.N, by=acct_date] # count occurrences by account date
a[order(-N)] # put in order
```

There seem to be a lot of accounts created yesterday suddenly tweeting this (a clue!). How many were made in the last month?
```{r}
a[acct_date>=ymd('2021-08-11')] # look at all the ones created in the past month

```

Back to the original, what's the spread on repeated creation dates?
```{r}
# get the quantiles for probabilities in the sequence 0-1 by 10% (.1)
quantile(a$N, probs = seq(0,1,by=.1))
```
80% of our subjects have creation dates that coincide with 3 or more others.

Let's look at that with a plot in visNetworks (it's interactive HTML - zoom in and out, select a node to highlight nearest neighbors):

```{r}
b<-a[N>4] # get a with N>4
edges<-b[,.(from=screen_name, to=as.character(acct_date))] # make a to-from edge frame from b
nodes<-data.table(id=unique(c(b$screen_name,as.character(b$acct_date))))
edges[,length:=.N, by=from] # count coincidents and label that "length"
nodes[id %in% edges$to, shape:='ellipse'] # make a column called shape and change the dates to an ellipse
nodes[id %in% edges$to, color:='red'] # do the same for color
nodes[,label:=id] # put a label on the nodes the same as the id
# plot it using an igraph layout with options to highlisht nearest neighbors
visNetwork(nodes=nodes, edges=edges, w=1280, h=1024) %>% visIgraphLayout(layout = 'layout_with_fr') %>% visOptions(highlightNearest = TRUE)

```


Now let's make a new set with the string distance between the names (there are more elegant ways to do this, but I'm doing this quickly):

```{r}
snames<-b$screen_name # get all the screen names from our b frame
m<-stringdistmatrix(snames, method = 'lv') # make a stringdist matrix showing the distance between names
mdt<-as.data.table(as.matrix(m)) # convert that to a data table
colnames(mdt)<-snames # set the column names to the actual strings
# create a new frame with the strings as a column (because the data.table doesn't keep row names)
sndt<-data.table(name=snames) 
mdt<-cbind(sndt,mdt) # column bind them
mmdt<-melt(mdt,id.vars = 'name') # fast "melt" the square matrix into a long table name to name to distance
mmdt<-mmdt[value>0] # take out the 0 distance ones (string to itself)
# make a new blank data table and keep only the shortest edges for each node 
dt<-data.table()
for (n in snames){
  d<-rbind(mmdt[name==n][value==min(value)],mmdt[variable==n][value==min(value)]) 
  dt<-rbind(dt,d)
  }
nedges<-dt[,.(from=name, to=variable, length=value)] # create new edges
nedges[,to:=as.character(to)] # Column names get converted to a "factor" class - convert them back
nnodes<-data.table(id=unique(c(nedges$to,nedges$from))) # make a new node frame with the unique to and from
# make an igraph network from the edge frame
net<-graph_from_data_frame(nedges)
cfg<-cluster_fast_greedy(as.undirected(net)) # use igraph fast,greedy clustering to cluster by modularity
groupnodes<-data.table(id=cfg$names, group=cfg$membership) # make a table of the results from clustering
nnodes<-merge(nnodes,groupnodes) # merge the tables together to add a group column
# plot as before
visNetwork(nodes=nnodes, edges=nedges, w=1280, h=1024) %>% visIgraphLayout(layout = 'layout_with_fr') %>% visOptions(highlightNearest = TRUE)

```

Put the two graphs together, but limit it to accounts created in the last month:
```{r}
d<-a[acct_date>=ymd('2021-08-10')] # make a new table with the last month's users
dnodes<-c(d$screen_name,as.character(d$acct_date)) # get a list of those users for filtering 
# bind our two previously created tables
alledges<-rbind(edges,nedges)
# pick the ones that were created in the last month
alledges<-alledges[from %in% dnodes & to %in% dnodes]
allnodes<-data.table(id=unique(c(alledges$from,alledges$to))) # done this a couple of times now

allnet<-graph_from_data_frame(alledges) # make the graph
ceb<-cluster_edge_betweenness(as.undirected(allnet)) # try a different clustering technique
cebdf<-data.table(id=ceb$names, group=ceb$membership) # get the results
allnodes<-merge(allnodes,cebdf) # add back the results
allnodes[id %in% as.character(a$acct_date), shape:='square'] # set a shape for the dates 
# plot that thing!
visNetwork(nodes=allnodes, edges=alledges, w=1280, h=1024) %>% visIgraphLayout(layout = 'layout_with_fr') %>% visOptions(highlightNearest = TRUE)
```

